---
title: "CapestoneEarthquake"
author: "Olga Vassilieva"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This package contains functions to visualise earthquakes from a dataset obtained
from the U.S. National Oceanographic and Atmospheric Administration (NOAA).
The functions are:
- data cleaning functions: eq_clean_data() and eq_location_clean
- geoms to visualise earthquakes magnitude and number of death by selected countries on the timeline: geom_timeline() and geom_timeline_label()
- functions mapping the earthquakes and label them eq_map() and eq_create_label()

## Obtain data

Download data from NOAA website, save it and load:

```{r, eval=FALSE}
library(readr)

NOAA_raw_data<-system.file("extdata","signif.txt",package="CapestoneEarthquake")
earthquake_track <- readr::read_tsv(NOAA_raw_data, col_names = TRUE)
as.data.frame(earthquake_track)

```
This function returns a dataframe "earthquake_track"

## Data cleaning

Before running visualisations clean the data using function eq_data_clean(). This function creates the data column by uniting the year, month, day and converts it 
to the Date class. LATITUDE and LONGITUDE columns are converted to numeric class and some other data (number of death, magnitude) are converting to numeric.

```{r, eval=FALSE}
library(dplyr)
library(magrittr)
 
earthquake_track_clean<-eq_clean_data(earthquake_track)

```

Next cleaning function eq_location_clean strips out the country name (including the colon) from the LOCATION_NAME column and converts names to title case (as opposed to all caps)

```{r, eval=FALSE}
library(dplyr)
library(stringr)
library(magrittr)
 
earthquake_track_clean<-eq_location_clean(earthquake_track_clean)

```


## Visualisation on the plot

Make a plot to visualize the times, the magnitudes and the number of deaths associated to earthquakes within certain countries, use the geom_timeline  with the `ggplot` function: 


```{r, eval=FALSE}
library(ggplot2)
library(grid)

earthquake_track_clean %>%
  dplyr::filter(COUNTRY %in% c('JAPAN', 'IRAN',YEAR > 2010)) %>%
  ggplot() +
  geom_timeline(aes(x = DATE, y = COUNTRY, color = TOTAL_DEATHS,
                    size = EQ_PRIMARY)) +
  scale_size_continuous(name = 'Richter scale value') +
  scale_color_continuous(name = '# of Deaths')+
  theme(legend.position="bottom", legend.box="horizontal")

```

In order to add labels to the plot use the geom_timelinelabel. It adds annotations to the earthquake data and an option to subset the n_max largest earthquakes by magnitude:

```{r, eval=FALSE}
library(ggplot2)
library(grid)
library(dplyr)

earthquake_track_clean%>%
  dplyr::filter(COUNTRY %in% c("JAPAN","IRAN","PERU") & lubridate::year(DATE) >= 2016) %>%
  ggplot() +
  geom_timeline(aes(x=DATE,y=COUNTRY,color=TOTAL_DEATHS,size=EQ_PRIMARY)) +
  geom_timeline_label(aes(label=LOCATION_NAME),n_max=3)+
  scale_size_continuous(name = 'Richter scale value') +
  scale_color_continuous(name = '# of Deaths')+
  theme(legend.position="bottom", legend.box="horizontal")

```

## Visualisation on the map

eq_map function helps to visualize earthquake epicentres of the certain region for the selected period. The column from the data can be chosen for annotation that will be in pop-up windows:

```{r, eval=FALSE}
eq_clean_data(earthquake_track) %>%
  dplyr::filter(COUNTRY == "MEXICO" & lubridate::year(DATE) >= 2000) %>% 
  eq_map(annot_col = "DATE")
```

To add more information to annotations use eq_create_lable function together with eq_map. This function adds the Location, Total deaths, and Magnitude of the earthquakes:

```{r, eval=FALSE}
eq_clean_data(earthquake_track) %>%
  dplyr::filter(COUNTRY == "IRAN" & lubridate::year(DATE) >= 2016) %>% 
  dplyr::mutate(popup_text = eq_create_label(.)) %>% 
  eq_map(annot_col = "popup_text")
```









